---
title: "ma350-lab5-RG-ML-SC"
author: "RG, Matt Lawson, and Sam Cavalli"
date: "3/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Git and GitHub

## 0

https://github.com/rmgreenstone/ma350-lab5-RG-ML-SC.git

# Finding Outliers Using Tukey’s Rule-of-Thumb

## Data


```{r}

x <- c(2.2, 7.8, -4.4, 0.0, -1.2, 3.9, 4.9, -5.7, -7.9, -4.9, 28.7, 4.9)
      
```


## 1



```{r}
quantile(x)
sort(x)
quant.4 <- quantile(x)[4]
quant.2 <- quantile(x)[2]
quant.4-quant.2
```

The first quartile is -4.525, the second quartile is 1.1 and the third quartile is 4.9. The interquartile range of x is Q3 - Q1 which is 9.425.I know these values are all correct based off arithmetic. 


## 2

```{r}
#' Take a vector containing a sample and return a vector with three components: the first quartile, the third quartile, and the interquartile range
#' @param X vector containing a sample
#'
#' @return a vector with the first quartile, the third quartile, and the interquartile range

quartiles <- function(X){

  the.quantiles <- quantile(X)
  quant.one <- quantile(X)[2]
  quant.three <- quantile(X)[4]
  inter <- quant.three-quant.one
  
   quart.vector<- c(quant.one, quant.three, inter)
  names(quart.vector) <- c("1st Quartile", "3rd Quartile", "Interquartile Range")
  
  return(quart.vector)
 
}
quartiles(x)
```



## 3


By inspection, it appears that 28.7 will be the only outlier.




## 4



```{r}
#' Take a vector containing a sample and find the outliers
#' @param X vector containing a sample
#'
#' @return a Boolean vector with TRUE at the elements corresponding to outliers and FALSE everywhere else 

tukey.outlier <- function(X){
  outlier <- vector(length = length(X))
  for (i in 1:length(X)) {
    if(X[i] < quartiles(X)[1]-1.5*quartiles(X)[3] || X[i] > quartiles(X)[2]+1.5*quartiles(X)[3]){
      outlier[i] <- TRUE
    } else {
      outlier[i] <- FALSE
    }
  }
  return(outlier)
}

tukey.outlier(x)
```

The function gives the correct output because 28.7 is the only outlier and is the only element in the vector that returned a TRUE.

## 5


```{r}
test.tukey.outlier <- function(fun = tukey.outlier){
  x <- c(2.2, 7.8, -4.4, 0.0, -1.2, 3.9, 4.9, -5.7, -7.9, -4.9, 28.7, 4.9)
  
  outliers.booleans <- c(FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, TRUE, FALSE)
  
  if(sum(fun(x)==outliers.booleans)==length(x)){
    return(TRUE)
  }else{
    return(FALSE)
  }
  
}
test.tukey.outlier(tukey.outlier)
```

The function test.tukey.outlier gives the correct answer when passed tukey.outlier. 

## 6

The data value that should be outliers in -x is -28.7.


## 7


```{r}
test.tukey.outlier <- function(fun = tukey.outlier){
  x <- c(2.2, 7.8, -4.4, 0.0, -1.2, 3.9, 4.9, -5.7, -7.9, -4.9, 28.7, 4.9)
  
negative.x <- c(-2.2, -7.8, 4.4, 0.0, 1.2, -3.9, -4.9, 5.7, 7.9, 4.9, -28.7, -4.9)
  

  outliers.booleans <- c(FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, TRUE, FALSE)
  
  out.result <- rep(NA,2)
  
  if(sum(fun(x)==outliers.booleans)==length(x)){
    out.result[1] <- TRUE
  }else{
    out.result[1] <- FALSE
  }
  
 if(sum(fun(negative.x)==outliers.booleans)==length(negative.x)){
   out.result[2] <- TRUE
 }else{
   out.result[2] <- FALSE
 }
 return(out.result) 
}

```



## 8


```{r}
test.tukey.outlier(tukey.outlier)
```

Since we get two TRUE's this means that tukey.outlier() passes the new set of tests.


## 9


The data value that should be an outlier in 0.1x is 2.87.



## 10



```{r}
test.tukey.outlier <- function(fun = tukey.outlier){
  x <- c(2.2, 7.8, -4.4, 0.0, -1.2, 3.9, 4.9, -5.7, -7.9, -4.9, 28.7, 4.9)
  
negative.x <- c(-2.2, -7.8, 4.4, 0.0, 1.2, -3.9, -4.9, 5.7, 7.9, 4.9, -28.7, -4.9)

tenth.x <- c(.22, .78, -.44, 0.0, -.12, .39, .49, -.59, -.79, -.49, 2.87, .49)
  

  outliers.booleans <- c(FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, TRUE, FALSE)
  
  out.result <- rep(NA,3)
  
  if(sum(fun(x)==outliers.booleans)==length(x)){
    out.result[1] <- TRUE
  }else{
    out.result[1] <- FALSE
  }
  
 if(sum(fun(negative.x)==outliers.booleans)==length(negative.x)){
   out.result[2] <- TRUE
 }else{
   out.result[2] <- FALSE
 }
  
   if(sum(fun(tenth.x)==outliers.booleans)==length(tenth.x)){
   out.result[3] <- TRUE
 }else{
   out.result[3] <- FALSE
 }
 return(out.result) 
}
```



## 11



```{r}
test.tukey.outlier(tukey.outlier)
```





